Finding ARG1 of Partitive Nouns in NomBank with DistilBERT

### 2022-05-05 src/paper.tex
Update

% Standard package includes
\usepackage{times}
\usepackage{latexsym}
\usepackage{float}
\usepackage{graphicx}

@@

@@

% Author information can be set in various styles:
% For several authors from the same institution:
% \author{Author 1 \and ... \and Author n \\
%         Address line \\ ... \\ Address line}
% if the names do not fit well on one line use
%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
% For authors from different institutions:
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \And  ... \And
%         Author n \\ Address line \\ ... \\ Address line}
% To start a seperate ``row'' of authors use \AND, as in
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \AND
%         Author 2 \\ Address line \\ ... \\ Address line \And
%         Author 3 \\ Address line \\ ... \\ Address line}

@@

More recently, the NLP community has seen huge popularity in heavily adopting pretraining-based langauge models in attempt to transfer knowledge learned from large corpus into more specific tasks. BERT (Devlin et al., 2018) is a large-size language representation model trained on a large corpus of English text that can achieve state-of-the-art results on a variety of natural language processing tasks. BERT came a decade later than the release of NomBank and there's been little previous work using it to conduct NomBank-based tasks and related experiments. However, as powerful as BERT is, its large parameters size makes it significantly more computationally expensive to train and use. On the other hand, DistilBERT is a distilled student model of BERT that retains 97\% of BERT's language understanding capabilities while being 40\% smaller and 60\% faster. This makes it a great candidate to train when facing with limited resources.

@@

In addition to the token classification approaches above, this paper also experiments with question-answer task, transforming the original ARG1 finding task into the task of make the model answer the question "What is the ARG1 of this sentence?". The advantage of this adaptation is that the model would stably output exactly one ARG1 for each sentence whereas token classification could give each sentence 0 to N ARG1s.

@@

@@

In practice, this pipeline has the tendency of over predicting ARG1 labels. It predicts on a per-token basis, making classification of whether the token is an ARG1 or not on every token, thus possibly producing 0 to multiple ARG1s in one sentence. It is shown by our experiments that the deep models are more common to produce more than 1 ARG1s in one sentence. The ground truth, however, is one and only one ARG1 per sentence. A treatment of this inconsistency is to always force one ARG1 by the logits output of the deep neural network. We assign the token with the largest logit value in the output vector for the whole sentence to be the ARG1.

@@

In a sentence in the \% dataset, the PRED labels \% symbol, the ARG1 is the partitive noun that \% symbol refers to, and the SUPPORT marks the word connecting the ARG1 and the PRED. It is guaranteed that there is one and only one ARG1 in each sentence of the dataset.

@@

All deep models used in this paper are implemented in PyTorch and with HuggingFace transformers libraries. The pre-trained models are fine-tuned on a single NVIDIA P100 (16GB) GPU on the Google Colab platform. The deep models are fine-tuned for 10 epochs with the learning rate of \verb|2e-5| and weight decay of \verb|0.01|. For token classification tasks, the batch size is 64. For question-answering tasks, the input would be bigger with the question prompt, so the batch size is set to 32 to be able to fit into the GPU's RAM.

@@

@@

\begin{table*}
  \centering
  \begin{tabular}{c||c|c|c}
    \hline
    \textbf{Model}       & \textbf{Precision} & \textbf{Recall} & \textbf{F1-Score} \\
    \hline
    MaxEnt Baseline      & {55.33\%}          & {36.02\%}       & {43.64\%}         \\
    DistilBERT           & {80.49\%}          & {78.43\%}       & {79.45\%}         \\
    DistilBERT (POS+BIO) & {81.53\%}          & {76.94\%}       & {79.16\%}         \\
    \hline
  \end{tabular}
  \caption{Precision, Recall and F1-Score on the partitive dataset}
  \label{tab:results-on-partitive-dataset}
\end{table*}

@@

As shown by Table \ref{tab:results-on-first-dataset}, the MaxEnt baseline achieved an F1 score of \verb|0.6619|. The DistilBERT models perform much better than the MaxEnt Baseline model on the \% dataset generally even when given less features, achieving an F1 score of 0.9184 for the basic pipeline,  \verb|0.9225| for the feature-enhanced pipeline,  \verb|0.92| for the QA pipeline and  \verb|0.9267| for the ONE ARG1 pipeline. The performance differences between different DistilBERT-based pipelines, however, are not significant. POS tags and NG-BIO tags likely don't have much effect on the performance of the model.

As for the partitive dataset, it is also shown by Table \ref{tab:results-on-partitive-dataset} that the DistilBERT model performs better than the MaxEnt Baseline model that achieved an F1 score of  \verb|0.4364|. The basic pipeline achieved an F1 score of  \verb|0.7945|. It is also observed that the POS tags and NG-BIO tags features didn't have significant impact on the performance of the model, achieving an F1 score of only  \verb|0.7916|.

@@

\section{Future Work}

@@

This document has been adapted
by Steven Bethard, Ryan Cotterell and Rui Yan
from the instructions for earlier ACL and NAACL proceedings, including those for
ACL 2019 by Douwe Kiela and Ivan Vuli\'{c},
NAACL 2019 by Stephanie Lukin and Alla Roskovskaya,
ACL 2018 by Shay Cohen, Kevin Gimpel, and Wei Lu,
NAACL 2018 by Margaret Mitchell and Stephanie Lukin,
Bib\TeX{} suggestions for (NA)ACL 2017/2018 from Jason Eisner,
ACL 2017 by Dan Gildea and Min-Yen Kan,
NAACL 2017 by Margaret Mitchell,
ACL 2012 by Maggie Li and Michael White,
ACL 2010 by Jing-Shin Chang and Philipp Koehn,
ACL 2008 by Johanna D. Moore, Simone Teufel, James Allan, and Sadaoki Furui,
ACL 2005 by Hwee Tou Ng and Kemal Oflazer,
ACL 2002 by Eugene Charniak and Dekang Lin,
and earlier ACL and EACL formats written by several people, including
John Chen, Henry S. Thompson and Donald Walker.
Additional elements were taken from the formatting instructions of the \emph{International Joint Conference on Artificial Intelligence} and the \emph{Conference on Computer Vision and Pattern Recognition}.

>>>>

% Standard package includes
\usepackage{lastpage}
\usepackage{fancyhdr}
\usepackage{times}
\usepackage{latexsym}
\usepackage{float}
\usepackage{graphicx}

@@

\pagestyle{fancy}
\fancyhead{}
\renewcommand{\headrulewidth}{0pt}
\fancyfoot[C]{Page \thepage \hspace{1pt} of \pageref{LastPage}}

@@

% Author information can be set in various styles:
% For several authors from the same institution:
% \author{Author 1 \and ... \and Author n \\
%         Address line \\ ... \\ Address line}
% if the names do not fit well on one line use
%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
% For authors from different institutions:
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \And  ... \And
%         Author n \\ Address line \\ ... \\ Address line}
% To start a separate ``row'' of authors use \AND, as in
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \AND
%         Author 2 \\ Address line \\ ... \\ Address line \And
%         Author 3 \\ Address line \\ ... \\ Address line}

@@

More recently, the NLP community has seen huge popularity in heavily adopting pretraining-based language models in attempt to transfer knowledge learned from large corpus into more specific tasks. BERT (Devlin et al., 2018) is a large-size language representation model trained on a large corpus of English text that can achieve state-of-the-art results on a variety of natural language processing tasks. BERT came a decade later than the release of NomBank and there's been little previous work using it to conduct NomBank-based tasks and related experiments. However, as powerful as BERT is, its large parameters size makes it significantly more computationally expensive to train and use. On the other hand, DistilBERT is a distilled student model of BERT that retains 97\% of BERT's language understanding capabilities while being 40\% smaller and 60\% faster. This makes it a great candidate to train when facing with limited resources.

@@

In addition to the token classification approaches above, this paper also experiments with question-answer task, reframing the original ARG1 finding task into the task of make the model answer the question "What is the ARG1 of this sentence?". The advantage of this adaptation is that the model would stably output exactly one ARG1 for each sentence whereas token classification could give each sentence 0 to N ARG1s.

@@

We developed several pipelines for finding ARG1 of partitive nouns in NomBank. One pipeline utilizes maximum entropy classifiers, the others utilizes the pre-trained DistilBERT model.

@@

In practice, this pipeline has the tendency of over generating ARG1 labels. It predicts on a per-token basis, making classification of whether the token is an ARG1 or not on every token, thus possibly producing 0 to multiple ARG1s in one sentence. It is shown by our experiments that the deep models are more common to produce more than 1 ARG1s in one sentence. The ground truth, however, is one and only one ARG1 per sentence. A treatment of this inconsistency is to always force one ARG1 by the logits output of the deep neural network. We assign the token with the largest logit value in the output vector for the whole sentence to be the ARG1.

@@

In a sentence in the \% dataset, the PRED labels \% symbol, the ARG1 is the partitive noun that \% symbol refers to, and the SUPPORT marks the word connecting the ARG1 and the PRED. It is guaranteed that there is one and only one ARG1 in each sentence of the \% dataset.

@@

All deep models used in this paper are implemented in PyTorch and with HuggingFace transformers libraries. The pre-trained models are fine-tuned on a single NVIDIA P100 (16GB) GPU on the Google Colab platform. The baseline maximum entropy classifier implementation we used is \verb|opennlp.maxent| by Baldridge et al. The deep models are fine-tuned for 10 epochs with the learning rate of \verb|2e-5| and weight decay of \verb|0.01|. For token classification tasks, the batch size is 64. For question-answering tasks, the input would be bigger with the question prompt, so the batch size is set to 32 to be able to fit into the GPU's RAM.

@@

\begin{table*}[t]
  \centering
  \begin{tabular}{c||c|c|c}
    \hline
    \textbf{Model}       & \textbf{Precision} & \textbf{Recall} & \textbf{F1-Score} \\
    \hline
    MaxEnt Baseline      & {55.33\%}          & {36.02\%}       & {43.64\%}         \\
    DistilBERT           & {80.49\%}          & {78.43\%}       & {79.45\%}         \\
    DistilBERT (POS+BIO) & {81.53\%}          & {76.94\%}       & {79.16\%}         \\
    \hline
  \end{tabular}
  \caption{Precision, Recall and F1-Score on the partitive dataset}
  \label{tab:results-on-partitive-dataset}
\end{table*}

@@

@@

\subsection{\% Dataset}
\label{sec:results-first-dataset}

As shown by Table \ref{tab:results-on-first-dataset}, the MaxEnt baseline achieved an F1 score of \verb|0.6619|. The DistilBERT models perform much better than the MaxEnt Baseline model on the \% dataset generally even when given less features, achieving an F1 score of 0.9184 for the basic pipeline,  \verb|0.9225| for the feature-enhanced pipeline,  \verb|0.92| for the QA pipeline and  \verb|0.9267| for the ONE ARG1 pipeline. However, The performance differences between different DistilBERT-based pipelines are very minor and can possibly be counted as experimental error. The introduction of additional POS tags and NG-BIO tags features likely don't seem to improve the model's performance in any statistically significant way.

A major advantage of DistilBERT pipelines over the baseline is the limited number of features needed. The basic and the QA pipelines only require the text input itself and feature-enhanced one requires the text input, the POS tags and the NG-BIO tags.

By a closer look to where the model struggles to produce correct predictions, Figure \ref{fig:prediction-one-sample} being one of the examples, we can see that the model is trying to predict more than one ARG1s in the sentence. In the given example, the ground truth is only \verb|tons| and the model is trying to predict \verb|tons| and also \verb|steel|. Though being over generating ARG1s, this is actually a very sensible prediction since "tons of steel" is indeed a noun group that the \% symbol is referring to. And our solutions to this over generating problem (QA task and forcing one ARG1) didn't quite improve the performance. In our experiment, we found that, when faced with multiple ARG1 candidates in a noun group, the model tends to choose the head noun as the ARG1, which in this case is \verb|steel| instead of the ground truth \verb|tons|.

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{assets/prediction-one-sample.png}
  \caption{One over-generating prediction sample for DistilBERT}
  \label{fig:prediction-one-sample}
\end{figure}

As for the partitive dataset, it is also shown by Table \ref{tab:results-on-partitive-dataset} that the DistilBERT model performs better than the MaxEnt Baseline model that achieved an F1 score of \verb|0.4364|. The basic pipeline achieved an F1 score of  \verb|0.7945|. It is also observed that the POS tags and NG-BIO tags features didn't have statistically significant improvement on the performance of the model, achieving an F1 score of only \verb|0.7916|. When compared to the \% dataset, the decrease in performance is expected since the partitive dataset has more distractions and noise. The tendency of over generating ARG1s is also observed for DistilBERT models in the partitive dataset, and can not easily be improved by forcing one ARG1, which is consistent with the observations made for the result on \% dataset shown in \ref{sec:results-first-dataset}.

@@

This paper has discussed the semantic role labeling task of finding ARG1 of partitive nouns in NomBank. And we have experimented with multiple ARG1 finding models and pipelines on \% dataset and partitive dataset from the NomBank corpus, taking advantages of the state-of-the-art DistilBERT model, then successfully developed a SRL system that can achieve an F1 score of \verb|0.9267| on the test set of the \% dataset and an F1 score of \verb|0.7945| on combined dataset of both \% dataset and partitive dataset. In comparisons with the baseline model, pre-trained DistilBERT models have shown their great advantage of text representation in this specific NLP task.

By the results of our experiments on feature-enhanced pipelines, we also show that the DistilBERT model cannot be significantly improved by adding POS tags and NG-BIO tags features. This is consistently shown by our experiments on both \% dataset and partitive dataset. Besides, for the \% dataset, question-answering variant and forcing one ARG1 variant of DistilBERT models have not shown much significant improvement on the performance of the model either. However, the QA approach is indeed a novel perspective to look at the task of finding ARG1, reframing a SRL task into a QA task which is a more open and flexible task. The flexibility comes with the free choice of prompt question - a prompt question can be any text. Therefore, what could possibly be further investigated is the question prompt we use to ask the model to find ARG1.

\section{Discussion and Future Work}

@@
